Commutative Character Sheaves on smooth group schemes over finite fields

### 2015-12-01 followup.tex
(re)starting the application to types

\title[Commutative Character Sheaves]{Commutative Character Sheaves}
%with an application to geometric types for supercuspidal representations
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, AB, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from the data. 
Let us refer to the types arising in this manner as {\it Yu types}.
In this section we show that every Yu type is geometrizable, in a sense which will be made clear below.

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
In Section~\ref{ssec:review} we recall properties of $(\,^\circ K,\,^\circ \rho)$ which we allow us, in Section~\ref{ssec:geometrization}, to find a pair $(G, \mathcal{F})$ where $G$ is a smooth group scheme over $\Fq$ and $\mathcal{F}$ is a virtual constructible sheaf $\ell$-adic sheaf on $G$ from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

@@

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item[(0)] a compact group $\,^\circ K$;
\item[(1)] compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item[(2)] a continuous representation $\varpi_0$ of $J^0$, inflated from ...
\item[(3)] a quasicharacter $\varphi_i : J^i \to \CC^\times$, for $i=0, \ldots d$.
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item[(4)] a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item[(5)] a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item[(6)] an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item[(7)] a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item[(8)] a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group;
\item[(9)] an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
%\end{enumerate}
\item[(10)]The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$ in the sense of \cite{yu:construction}, so, together with the quotient $J^{i+1} \to H^{i+1}$, it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} 
%& (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} 
& \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
%
From this information, Yu constructs 
%\begin{enumerate}
\item[(11)] a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$,
constructed as follows.
%\end{enumerate}
%We will need some more detail about the construction of the representations $\,^\circ \rho^i$, so we review it below.
%
 We will need some detail about this construction, so we will review it here.
%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
This determines the Weil representation $W_{i+1}$ on the same representation space as the Heisenberg representation of $V_{i+1}^\sharp$.
Let $W_{i+1} \ltimes \omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_{i+1}} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& (J^0 \cdots J^i) & 
\end{tikzcd}
\]
%
Now $h_1^*(\omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0\otimes \varphi_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0\otimes \varphi_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:01a}*{}, the representation 
$h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.  
%
Likewise, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.  
%
Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$. 
%
The character of $\,^\circ \rho_d$ on $\,^\circ K = J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
for $k_i\in J^i$.

>>>>

\title[Commutative Character Sheaves]{Commutative Character Sheaves on smooth group schemes over finite fields}
%with an application to geometric types for supercuspidal representations
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, AB, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from the data. 
Let us refer to the types arising in this manner as {\it Yu types}.
In this section we show that every Yu type is geometrizable.
% in a sense which will be made clear below.

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
In Section~\ref{ssec:review} we recall properties of $(\,^\circ K,\,^\circ \rho)$ which we allow us, in Section~\ref{ssec:geometrization}, to find a pair $(G, \mathcal{F})$ where $G$ is a group scheme over $\Fq$ and $\mathcal{F}$ is a virtual constructible sheaf $\ell$-adic sheaf on $G$ from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

@@

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item[(0)] a compact group $\,^\circ K$;
\item[(1)] compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item[(2)] a continuous representation $\varpi_0$ of $J^0$, inflated from ...
\item[(3)] a quasicharacter $\varphi_i : J^0\cdots J^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item[(4)] a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item[(5)] a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item[(6)] an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item[(7)] a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item[(8)] a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group;
\item[(9)] an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
%\end{enumerate}
\item[(10)]The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$ in the sense of \cite{yu:construction}, so, together with the quotient $J^{i+1} \to H^{i+1}$, it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} 
%& (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} 
& \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
\end{enumerate}

From this information, Yu constructs 
a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$,
constructed as follows.
%\end{enumerate}
%We will need some more detail about the construction of the representations $\,^\circ \rho^i$, so we review it below.
%
We will need some detail about this construction,   so we review it here.

%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
This determines the Weil representation $W_{i+1}$ on the same representation space as the Heisenberg representation of $V_{i+1}^\sharp$.
Let $W_{i+1} \ltimes \omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_{i+1}} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& J^0 \cdots J^i & 
\end{tikzcd}
\]

Now $h_1^*(W_1\ltimes \omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:01a}*{}, the representation 
$h_1^*(\omega_1) \otimes \varphi_1 \otimes \inf(\sigma_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \varphi_2 \otimes\inf(\sigma_0\otimes)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.  
%
Likewise, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.  
%
Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$. 
%
The character of $\,^\circ \rho_d$ on $\,^\circ K = J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
for $k_i\in J^i$.
